---
title: Problem Set 1 - Q1 - Analysis on Education Level and Income for Those who Earn
  less than $1000000
author: "Chenyang Huan 1004728459"
date: "Sep.30.2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Question 1
# Part a 

```{r,include=FALSE,echo=FALSE}
#install.packages("opendatatoronto") #; or 
#install.packages("devtools")
#devtools::install_github("sharlagelfand/opendatatoronto") #; or
#devtools::install_github("hodgettsp/cesR")

library(ggplot2)
library(cesR)
library(tidyverse)
library(dplyr)
get_ces("ces2019_web")
colnames(ces2019_web)

```


I choose the 2019 Canadian Election Study - Online Survey since this is the most
current data that we could use to understand Canadian's political behavior and 
attitudes. Canadian society and political life could be revealed from this 
dataset. I will be focusing on the relationship between level of education and 
household income.The data set includes 37822 observations which is enough to 
make accurate analysis. 

# Part b


```{r,echo=FALSE}
ces2019_web1<-ces2019_web
unique(ces2019_web1$cps19_education)
ces2019_web1<-ces2019_web %>% filter(cps19_income_number < 1000000 &cps19_income_number>500)
ces2019_web1$cps19_education1<-as.character(ces2019_web1$cps19_education)
ces2019_web1 %>% 
  ggplot(aes(x=cps19_education1,y=cps19_income_number,fill=cps19_education1))+geom_boxplot()+
  labs(title="Boxplot of Income by Education Level",caption="2019 Canadian Election Stury - Online Survey dataset",
       x="Education Level",y="Total Household Income Before Taxes in 2018($)")+theme_classic() +theme(legend.position="")
                                                            

```

There are lots of outliers in every education level, and they are all right
skewed. "12" are people who prefer not to say, beside those people, we could see
a steady pattern of increasing in income with the growth of education 
level. For people have elementary school as their highest education level, there
is an approximately $25000 upper limit. The median income for those who have 
completed college are close to the median income of people completed some 
university.


# Part c
```{r,include=FALSE}
options(scipen=999)
ces2019_web1<-ces2019_web
summary(ces2019_web1$cps19_income_number)
ces2019_web1<-ces2019_web1 %>% filter(cps19_income_number < 1000000 &cps19_income_number>500)
unique(ces2019_web1$cps19_education)
ces2019_web1$cps19_education1<-as.character(ces2019_web1$cps19_education)
unique(ces2019_web1$cps19_education1)

```
```{r,echo=FALSE}
ces2019_web1 %>% 
  mutate(cps19_education=as.character(cps19_education1)) %>% 
  mutate(cps19_education1=case_when(cps19_education1=="1"~"No schooling",
                                    cps19_education1=="2"~"Elementary School",
                                    cps19_education1=="3"~"Elementary School",
                                    cps19_education1=="4"~"High School",
                                    cps19_education1=="5"~"High School",
                                    cps19_education1=="6"~"College",
                                    cps19_education1=="7"~"College",
                                    cps19_education1=="8"~"University",
                                    cps19_education1=="9"~"University",
                                    cps19_education1=="10"~"Master",
                                    cps19_education1=="11"~"Doctor",
                                    cps19_education1=="12"~"Don't know/Prefer not to answer"))  %>%        
  filter(!is.na(cps19_education1)&cps19_education1!="") %>% 
  group_by(cps19_education1) %>% 
  summarise(minimum=min(cps19_income_number,na.rm = T),
            mean=mean(cps19_income_number,na.rm = T),
            median=median(cps19_income_number,na.rm = T),
            maximum=max(cps19_income_number,na.rm = T),
            std=sd(cps19_income_number,na.rm = T))
            
```



Based on the summary of income, the mean and maximum are abnormally large, so I 
treat the ones that are greater than $100000 as outlier and filter them out. 
According to Fraser Institute News Release, Canadians who earn more than $96,000
are in the top 10%, so it's fair to set the upper bound of income as $1,000,000.
Notice that the minimum wage is zero for all education levels, a $500 lower 
bound was set to drop meaningless data.
Initially the educational levels were divided into 12 groups and labeled 
respectively. In order to reduce the number of rows, I renamed each group and 
combined similar groups into one category.Surprisingly people who haven't attend
school earn more than ones who went to high school. The median for people who 
have completed high school and the ones who had no schooling are $50000 and 
$71168 respectively. People who didn't attend post-secondary schools couldn't
get more than $250000 income. People completed college could get a maximum of 
\$999000  while the maximum income for a doctor is $942600.


# Part d

We investigate on the data of household income before tax and level of 
education and analyze the data using "summurize" by different education level
groups.
It's commonly known that the more we learn, the more we earn. By doing the 
analysis on education levels and income, it's true that with a higher degree we
could be offered with more salary. Gaining more in the future could be the 
motivation of learning hard during school. Studying isn't the only way to earn 
more and improve living standards, but it's a relatively conservative way 
with high payback.

Weakness of the analysis: There are other lurking variables that would affect
household income. For example, partner's income, years of working experiences.
Also, missing data was filtered out.

Next steps for investigating could be making a multivariate model and includes
those lurking variables.

# Part e

Bibliography:

1. Stephenson, Laura B; Harell, Allison; Rubenson, Daniel; Loewen, Peter John, 
2020, '2019 Canadian Election Study - Online Survey',https://doi.org/10.7910/DVN/DUS88V, Harvard Dataverse, V1

2. R Core Team (2020). R: A language and environment for statistical computing. 
R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

3. Hadley Wickham, Romain François, Lionel Henry and Kirill Müller (2020). 
dplyr: A Grammar of Data  Manipulation. R package version 1.0.2. https://CRAN.R-project.org/package=dplyr
  
4. H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New 
York, 2016.

5. Hadley Wickham, Jim Hester and Winston Chang (2020). devtools: Tools to Make
  Developing R Packages Easier. R package version 2.3.2.
  https://CRAN.R-project.org/package=devtools
